Understanding Structure and Selectivity in Nanostructured Lyotropic Liquid Crystal Membranes using Molecular Simulations

### 2020-03-08 Ben_Manuscripts/thesis/intro.tex
Some comments on intro.

  %BJC: limiting to pressure driven aqueous membrane separations (so I don't have to get into electrodialysis, gas separations, dialysis etc.)
  Pressure driven membrane processes have become an increasingly useful tool for
  performing aqueous separations.
  \begin{itemize}
    \item Untreated water can be a very complex solution with particles ranging
    in size from microns down to nanometers
    \item Sediment, bacteria, algae, proteins, small organic molecules, and ions
    are all common components of aqueous streams.
  \end{itemize}

  % TFC membranes review -- "Composite reverse osmosis and nanofiltration membranes"
  \noindent Membrane design is completely dependent on the target particle separation. 
  \begin{itemize}
    \item Membranes are classified based on the size of the particle they reject.
    \item Microfiltration membranes contain pores with diameters ranging from 100-10000 nm.
    They can separate large particles like bacteria and protozoa.~\cite{ma_functionalized_2014}
    \item Ultrafiltration membranes have pore diameters of about 5-500 nm and are useful for the 
    separation of sugars, proteins, viruses and colloidal materials.~\cite{yanful_ultrafiltration_2009}
    \item Nanofiltration have pores on the order of 1 nm in diameter and can be used for the separation
    of small organic molecules and charged species.~\cite{van_der_bruggen_review_2003}
    \item Reverse osmosis membranes are dense amorphous polymers with no explicit pores. Their
    dense polymer architecture rejects all solutes to a high degree except for water making
    them useful for separating hydrated ions from water.~\cite{warsinger_review_2018}
    \item We have summarized the uses of different classes of membrane 
    technologies in Figure~\ref{fig:size_regimes}.
    \item To deal with complex aqueous streams, and to prevent excessive membrane fouling, 
    these technologies are often used in series, removing larger particles first.~\cite{goosen_fouling_2005}
  \end{itemize}

@@

  \noindent Membrane permeation is the result of a chemical potential gradient, $\frac{d\mu_i}{dx}$.
  \begin{itemize}
	  \item The flux of component $i$ is proportional to this gradient:
	  \begin{equation}
	    J_i = -L_i \frac{d\mu_i}{dx}
	  \end{equation}
	  where $L_i$ is a coefficient of proportionality.~\cite{wijmans_solution-diffusion_1995}
	  \item Chemical potential gradients can be induced by concentration, pressure, 
	  temperature and electromotive forces.
	  \item We will limit our discussion to concentration and pressure gradients.
  \end{itemize}

  % This follows the discussion of werber_material_2016
  For pressure driven membrane processes, water flux through both porous and 
  amorphous membranes can be described by
  \begin{equation}
  J_w = A(\Delta P - \Delta \pi_m)
  \end{equation}
  where $J_w$ is volumetric water flux, $A$ is the water permeability coefficient,
  $\Delta P$ is the applied hydraulic pressure and $\Delta \pi_m$ is the 
  trans-membrane osmotic pressure difference. The water permeability coefficient
  is dependent on how one models flow through the membrane.

  In porous membrane architectures, water flux is modeled as laminar flow through 
  cylindrical pores.
  \begin{itemize}
    \item The derivation follows Hagen-Poiseuille flow with inclusion of morphological
    characteristics:
    \begin{equation}
      A = \frac{\varepsilon r_p^2}{8\mu\delta_m}
    \end{equation}
    where $\varepsilon$ is the surface porosity, $r_p$ the pore radius, $\delta_m$ the
    membrane thickness, and $\mu$ the solution viscosity.
    \item Separation is typically characterized in terms of solute rejection, $R$. 
    \begin{equation}
      R = 1 - \frac{c_p}{c_f}
    \end{equation}
    where $c_p$ and $c_f$ are the concentrations of the rejected species in
    the permeate and feed respectively.
    \item Solutes with radii smaller than the pore will be rejected. A distribution of 
    pore sizes prevents perfect rejection.
  \end{itemize}

@@

  Although there has been considerable focus on creating membranes with high
  permeabilities, there is a case to be made that shifting the focus to membrane
  selectivity may offer a more effective route towards lowering the cost of 
  membrane separations.
  \begin{itemize}
    \item Increasing permeabilities beyond those already achieved may have 
    only a small effect on energy requirements and capital costs. 
    \item Cohen-Tanugi et al. calculated that tripling membrane permeability
    could reduce energy consumption up to 15\% in a seawater reverse osmosis plant.~\cite{cohen-tanugi_quantifying_2014}
    \item Perhaps the biggest detriment to energy consumption is the practical
	need for multiple membrane passes in order to achieve a desired purity.
    \item Additionally, many membranes are not suited for high purity separations
    of small neutral species which necessitates chemical and capital costs for
    post-treatment
    \item Addressing these problems with highly selective separations may lead to 
    more gains.
  \end{itemize}

@@

  Selective separations are useful in a wide range of applications.

@@

  Amorphous thin film composite (TFC) RO membranes are the current industry standard for
  high purity separations.
  \begin{itemize}  
    \item Desalination, concentration
    \item Typically you have a porous support layer for mechanical strength and 
    very thin active layer where the separations occur.
    \item Usually only let water through. Usually have to remineralize the
    water to make it drinkable.
  \end{itemize}

@@

  \subsection{Standard Nanofiltration Membranes}

  Unlike RO membranes, NF membranes have explicitly defined pores. An
  ideal NF membrane should have densely packed, uniform-sized and
  non-tortuous pores. This combination has been very difficult to realize.

@@

  Carbon nanotubes (CNTs) have shown promise as aqueous separations membranes
  due to unprecedentedly fast water transport.\cite{humplik_nanostructured_2011,hummer_water_2001}
  \begin{itemize}
	\item Practically, dispersing and aligning CNTs into a polymer matrix is extremely
    difficult because they tend to agglomerate due to Van der Waals forces.~\cite{sahoo_polymer_2010}
    \item Some work has been to functionalize the carbon nanotubes in order to better incorporate them. % need reference
  \end{itemize}

@@

@@

  In this work we study LLC membranes as an alternative nanostructured material for 
  highly selective aqueous separations.

@@

  The H\textsubscript{II} phase is characterized by hexagonally packed, uniform-sized
  and straight pores.
  \begin{itemize}
    \item The hydrophilic head groups aggregate in the pore centers to create
    cylindrical aqueous channels. 
    \item These pore channels are lined with the chemical functionality of the LLC monomers
    and have the potential to interact with solutes in a chemically-dependent manner.
    \item In theory, this pore topology and geometry is ideal for high flux and 
    highly selective separations.
  \end{itemize}

@@

  Due to its more facile synthesis, there has been significantly more development
  of Q\textsubscript{I} phase-forming monomers.
  \begin{itemize}
    \item The first generation of Q\textsubscript{I} membranes was made by 
    the self-assembly of a gemini phosphonium monomer in water.~\cite{pindzola_cross-linked_2003}
    \item Hatakeyama et al. improved the industrial viability of Q\textsubscript{I} 
    phase membranes by using a gemini ammonium monomer which is both easier and cheaper
    to synthesize.~\cite{hatakeyama_nanoporous_2010}
    \item Free standing films of both first and second generation Q\textsubscript{I} phase 
    membranes cannot withstand high pressure. Because solution casting was ineffective, Zhou
    et. al used a hot-pressing method to make mechanically strong membranes which involves 
    heating the initial monomer mixture to 70\degree C and pressing it with 12 tons of force
    into a thick microporous, hydrophilic polymer support.~\cite{zhou_new_2007}
    \item The most recent generation of Q\textsubscript{I} membranes uses an imidazolium-based
    gemini LLC monomer which is capable of being solution cast into defect-free thin films on
    porous supports. This improvement resulted in a ten-fold increase in flux while retaining
    selectivities similar to earlier generations.~\cite{carter_glycerol-based_2012}
  \end{itemize}

@@

  Over the past 20 years, H\textsubscript{II}-phase LLC polymer membrane
  studies have been limited primarily to the Na-GA3C11 monomer with some
  characterization done after minor structural modifications. For example, Resel
  et al.~varied the length of the monomer tails and the counterion used and
  observed its effect on pore spacing \cite{resel_structural_2000}. In a later
  study of rejection performance, it was shown that membranes formed by
  cross-linked Na-GA3C11 in the H\textsubscript{II} phase cannot separate solutes
  less than 1.2 nm in diameter because the pores are too large
  \cite{zhou_supported_2005}. We do not yet understand how to controllably reduce
  the effective pore size or how to tune the chemical environment in the
  nanopores of this or related materials for small molecule separations. The only
  source of predictive modeling for LLC systems have been macroscopic models that
  likely do not adequately describe transport at these length scales
  \cite{hatakeyama_water_2011}. Modeling with molecular detail could provide
  sufficient information about the mechanisms and chemical features to better
  inform experimental design of similar nanostructured membranes.

@@

  Because there is relatively sparse coverage of these types of simulation systems in 
  the literature, we have built a detailed molecular model from the ground up. There 
  are four primary research questions that we will address in this work.
  \begin{enumerate}
    \item What is the nanoscopic structure?
    	\begin{itemize}
    	  \item Before we can try to understand the molecular mechanisms of solute
    	  transport, we need to ensure that we model the chemical environment
    	  within the nanopores in a way that is consistent with experiment.
    	  \item We use experimental structural data in order to validate our model.
    	\end{itemize}
    \item Which solute-membrane interactions have the greatest influence on transport rates?
    	\begin{itemize}
    	  \item After gaining a detailed picture of the nanopore structure, we can 
    	  feel confident that solutes in this system will experience the same 
    	  interactions which are present in a real system.
    	  \item We create independent systems for each of 20 small polar solutes 
    	  and observe transport mechanisms whose dominance is dependent on
    	  solute chemical functionality.
    	  \item We characterize three different trapping mechanisms which lead
    	  to subdiffusive transport behavior.
    	\end{itemize}
    \item Can we estimate experimentally-relevant macroscopic properties?
    	\begin{itemize}
    	  \item Using our qualitative understanding of the dominant trapping mechanisms,
    	  we develop stochastic time series models which we can use to mimic solute dynamic
    	  behavior on time scales orders of magnitude longer than our simulations. 
    	  \item We attempt to reproduce both quantitative and qualitative solute trajectory
    	  behavior on MD simulation-length timescales.
    	  \item We then show how we can use our most promsiing model in order to connect 
    	  microscopic transport to macroscopic flux and selectivity.
    	\end{itemize}
    \item How can we learn mechanisms with minimal human intervention?  %BJC: i.e. machine learning methods
    	\begin{itemize}
    	  \item Although our stochastic models show great promise, their development 
    	  require some qualitative and quantitative understanding of dominant transport
    	  mechanisms.
    	  \item In the final chapter, we use the infinite hidden Markov model in order
    	  to automatically detect and parameterize a unknown number of hidden dynamical modes
    	  exhibited by solute time series. 
    	  \item This more flexible approach allows us to both infer mechanisms based 
    	  on differences in dynamical behavior and generate stochastic trajectory realizations
    	  which we can use to predict flux and selectivity. 
    	\end{itemize}
  \end{enumerate}

>>>>

  %BJC: limiting to pressure driven aqueous membrane separations (so I don't have to get into electrodialysis, gas separations, dialysis etc.)
  Pressure driven membrane processes have become an increasingly useful tool for
  performing aqueous separations.
  \begin{itemize}
    \item Untreated water can be a very complex solution with particles ranging
    in size from microns down to nanometers
    \item Sediment, bacteria, algae, proteins, small organic molecules, and ions
    are all common components of aqueous streams.
%MRS: maybe talk about molecular separations are particularly hard (to set up the work you do).
  \end{itemize}

  % TFC membranes review -- "Composite reverse osmosis and nanofiltration membranes"
  \noindent Membrane design is completely dependent on the target particle separation. 
  \begin{itemize}
    \item Membranes are classified based on the size of the particle they reject.
    \item Microfiltration membranes contain pores with diameters ranging from 100-10000 nm.
    They can separate large particles like bacteria and protozoa.~\cite{ma_functionalized_2014}
    \item Ultrafiltration membranes have pore diameters of about 5-500 nm and are useful for the 
    separation of sugars, proteins, viruses and colloidal materials.~\cite{yanful_ultrafiltration_2009}
    \item Nanofiltration have pores on the order of 1 nm in diameter and can be used for the separation
    of small organic molecules and charged species.~\cite{van_der_bruggen_review_2003}
    %MRS: no explicit pores might not quite be right. . . because there are pores there, just not designed pores. tortuous Pores formed by collapsed polymer networks? 
    \item Reverse osmosis membranes are dense amorphous polymers with no explicit pores. Their
    dense polymer architecture rejects all solutes to a high degree except for water making
    them useful for separating hydrated ions from water.~\cite{warsinger_review_2018}
    \item We have summarized the uses of different classes of membrane 
    technologies in Figure~\ref{fig:size_regimes}.
    \item To deal with complex aqueous streams, and to prevent excessive membrane fouling, 
    these technologies are often used in series, removing larger particles first.~\cite{goosen_fouling_2005}
  \end{itemize}

@@

%MRS: provide more of a context for this discussion.  Why are your reviewing this theory here?  
%MRS: are you assuming solution/diffusion here? If so, state explicitly. 
%MRS: OK after reading through, I have a suggestion.  Maybe the theory section should to nearer to the end.  At the beginning you talk more about membrane structure and applications, and talk about more traditional membranes, and then pivot to new membranes. Maybe you should stick with structure/function until you talk about the solute motion, and how it's NOT simple diffusioin, then talk about standard theory, then talk about the the need for alternative theory, and then introduce the ideas you will describe in much more detail in the chapters corresponding to your later papers.  Does that make sense as a strategy?
  \noindent Membrane permeation is the result of a chemical potential gradient, $\frac{d\mu_i}{dx}$.
  \begin{itemize}
	  \item The flux of component $i$ is proportional to this gradient:
	  \begin{equation}
	    J_i = -L_i \frac{d\mu_i}{dx}
	  \end{equation}
	  where $L_i$ is a coefficient of proportionality.~\cite{wijmans_solution-diffusion_1995}
          %MRS2: mention what L depends on molecularly (even if uncalculatable)
	  \item Chemical potential gradients can be induced by concentration, pressure, 
	  temperature and electromotive forces.
	  \item We will limit our discussion to concentration and pressure gradients. %MRS: mention why?
  \end{itemize}

  %MRS: is there are reason you are talkling about just water flux here?
  % This follows the discussion of werber_material_2016
  For pressure driven membrane processes, water flux through both porous and 
  amorphous membranes can be described by
  \begin{equation}
  J_w = A(\Delta P - \Delta \pi_m)
  \end{equation}
  where $J_w$ is volumetric water flux, $A$ is the water permeability coefficient,
  $\Delta P$ is the applied hydraulic pressure and $\Delta \pi_m$ is the 
  trans-membrane osmotic pressure difference. The water permeability coefficient
  is dependent on how one models flow through the membrane.

  %MRS: be clearer on presenting the context for this theory.  Will your findings be in contradiction?
  %MRS: is this theory inadequate?  Will it be tested?
  In porous membrane architectures, water flux is modeled as laminar flow through 
  cylindrical pores.
  \begin{itemize}
    \item The derivation follows Hagen-Poiseuille flow with inclusion of morphological
    characteristics:
    \begin{equation}
      A = \frac{\varepsilon r_p^2}{8\mu\delta_m}
    \end{equation}
    where $\varepsilon$ is the surface porosity, $r_p$ the pore radius, $\delta_m$ the
    membrane thickness, and $\mu$ the solution viscosity.
    \item Separation is typically characterized in terms of solute rejection, $R$. 
    \begin{equation}
      R = 1 - \frac{c_p}{c_f}
    \end{equation}
    where $c_p$ and $c_f$ are the concentrations of the rejected species in
    the permeate and feed respectively.
    \item Solutes with radii smaller than the pore will be rejected. A distribution of 
    pore sizes prevents perfect rejection.
  \end{itemize}

@@

  Although there has been considerable focus on creating membranes with high
  permeabilities, there is a case to be made that shifting the focus to membrane
  selectivity may offer a more effective route towards lowering the cost of 
  membrane separations.
  %MRS: citation?
  \begin{itemize}
    \item Increasing permeabilities beyond those already achieved may have 
    only a small effect on energy requirements and capital costs. 
    \item Cohen-Tanugi et al. calculated that tripling membrane permeability
    could reduce energy consumption up to 15\% in a seawater reverse osmosis plant.~\cite{cohen-tanugi_quantifying_2014}
    \item Perhaps the biggest detriment to energy consumption is the practical
	need for multiple membrane passes in order to achieve a desired purity.
    \item Additionally, many membranes are not suited for high purity separations
    of small neutral species which necessitates chemical and capital costs for
    post-treatment
    \item Addressing these problems with highly selective separations may lead to 
    more gains.
    %MRS: math (or citations) for this statement?
 \end{itemize}

@@

%MRS: Maybe the applications be presented before theory?
  Selective separations are useful in a wide range of applications.

@@

  Amorphous thin film composite (TFC) RO membranes are the current industry standard for
  high purity separations.
  \begin{itemize}  
    \item Desalination, concentration
    \item Typically you have a porous support layer for mechanical strength and 
    very thin active layer where the 
%MRS: added
molecular/nanoscale
separations occur.
    \item Usually only let water through. Usually have to remineralize the
    water to make it drinkable.
  \end{itemize}

@@

  \subsection{Standard Nanofiltration Membranes}
  %MRS: is explicitly defined the right word?  ``Created pores?''
  Unlike RO membranes, NF membranes have explicitly defined pores. An
  ideal NF membrane should have densely packed, uniform-sized and
  non-tortuous pores. This combination has been very difficult to realize.

@@

  Carbon nanotubes (CNTs) have shown promise as aqueous separations membranes
  due to unprecedentedly fast water transport.\cite{humplik_nanostructured_2011,hummer_water_2001}
  \begin{itemize}
	\item Practically, dispersing and aligning CNTs into a polymer matrix is extremely
    difficult because they tend to agglomerate due to van der Waals forces.~\cite{sahoo_polymer_2010}
    \item Some work has been to functionalize the carbon nanotubes in order to better incorporate them. % need reference
  \end{itemize}

@@

%MRS: probably could generalize to self-assmbled porous materials as
%the class in general, and the cross-linked LLC's as a particular exemplar.

@@

  %MRS: be more explicit what you mean by ``study''.
  In this work we study LLC membranes as an alternative nanostructured material for 
  highly selective aqueous separations.

@@

  The H\textsubscript{II} phase is characterized by hexagonally packed, uniform-sized
  and straight pores.
  \begin{itemize}
    \item The hydrophilic head groups aggregate in the pore centers to create
    cylindrical aqueous channels. 
    \item These pore channels are lined with the chemical functionality of the LLC monomers
    and have the potential to interact with solutes in a chemically-dependent manner.
    %MRS: could discuss concept that the effective pore size depends on the solute.
    \item In theory, this pore topology and geometry is ideal for high flux and 
    highly selective separations.
  \end{itemize}

@@

  %MRS: be more explicit that this is synthesis of usable amounts of
  %actual separations membranes, not the molecules themselves.
  Due to its more facile synthesis, there has been significantly more development
  of Q\textsubscript{I} phase-forming monomers.
  \begin{itemize}
    \item The first generation of Q\textsubscript{I} membranes was made by 
    the self-assembly of a gemini phosphonium monomer in water.~\cite{pindzola_cross-linked_2003}
    \item Hatakeyama et al. improved the industrial viability of Q\textsubscript{I} 
    phase membranes by using a gemini ammonium monomer which is both easier and cheaper
    to synthesize.~\cite{hatakeyama_nanoporous_2010}
    \item Free standing films of both first and second generation Q\textsubscript{I} phase 
    membranes cannot withstand high pressure. Because solution casting was ineffective, Zhou
    et. al used a hot-pressing method to make mechanically strong membranes which involves 
    heating the initial monomer mixture to 70\degree C and pressing it with 12 tons of force
    into a thick microporous, hydrophilic polymer support.~\cite{zhou_new_2007}
    \item The most recent generation of Q\textsubscript{I} membranes uses an imidazolium-based
    gemini LLC monomer which is capable of being solution cast into defect-free thin films on
    porous supports. This improvement resulted in a ten-fold increase in flux while retaining
    selectivities similar to earlier generations.~\cite{carter_glycerol-based_2012}
  \end{itemize}

@@

  Over the past 20 years, H\textsubscript{II}-phase LLC polymer membrane
  studies have been limited primarily to the Na-GA3C11 monomer with some
  characterization done after minor structural modifications. For example, Resel
  et al.~varied the length of the monomer tails and the counterion used and
  observed its effect on pore spacing \cite{resel_structural_2000}. In a later
  study of rejection performance, it was shown that membranes formed by
  cross-linked Na-GA3C11 in the H\textsubscript{II} phase cannot separate solutes
  less than 1.2 nm in diameter because the pores are too large
  %MRS: this above is interesting, since that seems somewhat inconsistent with the 
  \cite{zhou_supported_2005}. We do not yet understand how to controllably reduce
  the effective pore size or how to tune the chemical environment in the
  nanopores of this or related materials for small molecule separations. The only
  source of predictive modeling for LLC systems have been macroscopic models that
  likely do not adequately describe transport at these length scales
  \cite{hatakeyama_water_2011}. Modeling with molecular detail could provide
  sufficient information about the mechanisms and chemical features to better
  inform experimental design of similar nanostructured membranes.

@@

  Because there is relatively sparse coverage of these types of simulation systems in 
  the literature, we have built a detailed molecular model from the ground up. There 
  are four primary research questions that we will address in this work.
  \begin{enumerate}
    \item What is the nanoscopic structure?
    	\begin{itemize}
    	  \item Before we can try to understand the molecular mechanisms of solute
    	  transport, we need to ensure that we model the chemical environment
    	  within the nanopores in a way that is consistent with experiment.
    	  \item We use experimental structural data in order to validate our model.
    	\end{itemize}
    \item Which solute-membrane interactions have the greatest influence on transport rates?
    	\begin{itemize}
    	  \item After gaining a detailed picture of the nanopore structure, we can 
    	  feel confident that solutes in this system will experience the same 
    	  interactions which are present in a real system.
    	  \item We create independent systems for each of 20 small polar solutes 
    	  and observe transport mechanisms whose dominance is dependent on
    	  solute chemical functionality.
    	  \item We characterize three different trapping mechanisms which lead
    	  to subdiffusive transport behavior.
    	\end{itemize}
    %MRS: maybe talk more about the stochastic modeling itself; for example, could phrase it in terms of
    %how you can test mechanistic models to see if they are consistent with the timeseries, and then how those models
    %can be further extrapolated to physical observables even if the MD simulations cannot. 
    %MRS: this MIGHT include using the effective diffusion in a solution/diffusion model.
    \item Can we estimate experimentally-relevant macroscopic properties?
    	\begin{itemize}
    	  \item Using our qualitative understanding of the dominant trapping mechanisms,
    	  we develop stochastic time series models which we can use to mimic solute dynamic
    	  behavior on time scales orders of magnitude longer than our simulations. 
    	  \item We attempt to reproduce both quantitative and qualitative solute trajectory
    	  behavior on MD simulation-length timescales.
    	  \item We then show how we can use our most promsiing model in order to connect 
    	  microscopic transport to macroscopic flux and selectivity.
    	\end{itemize}
    %MRS: this might be overly ambitions.  Maybe something more like: How can the timeseries data inform us about the mechanism directly?
    \item How can we learn mechanisms with minimal human intervention?  %BJC: i.e. machine learning methods
    	\begin{itemize}
    	  \item Although our stochastic models show great promise, their development 
    	  require some qualitative and quantitative understanding of dominant transport
    	  mechanisms.
    	  \item In the final chapter, we use the infinite hidden Markov model in order
    	  to automatically detect and parameterize a unknown number of hidden dynamical modes
    	  exhibited by solute time series. 
    	  \item This more flexible approach allows us to both infer mechanisms based 
    	  on differences in dynamical behavior and generate stochastic trajectory realizations
    	  which we can use to predict flux and selectivity. 
    	\end{itemize}
  \end{enumerate}
